---
title: "Data-Management-Project"
author: "Alex Martinez & Serry Ezbidi"
format: html
---

```{r}
here::i_am("Data-Management-Project.Rproj")
library(here)
library(vroom)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(eurostat)
library(lubridate)
```

```{r}
crime <- get_eurostat("crim_just_ctz", time_format = "num", stringsAsFactors = TRUE)
manifesto <- vroom(("https://manifesto-project.wzb.eu/down/data/2024a/datasets/MPDataset_MPDS2024a.csv"))
ideology <- vroom(("https://eupoliticalbarometer.uc3m.es/api/ideologyDownload"))
```

Data Management

1- Link to GitHub
https://github.com/alexmar313/Data-Management-Project.git

2- Downloading links for data
https://eupoliticalbarometer.uc3m.es/api/ideologyDownload
https://manifesto-project.wzb.eu/down/data/2024a/datasets/MPDataset_MPDS2024a.csv
https://ec.europa.eu/eurostat/databrowser/view/crim_just_ctz/default/table?lang=en

3- Websites links from which data was extracted
https://eupoliticalbarometer.uc3m.es/dashboard/ideology
https://manifesto-project.wzb.eu/datasets
https://ec.europa.eu/eurostat/databrowser/view/crim_just_ctz/default/table?lang=en

4- Data description
Eurostat Crime and Criminal Justice Dataset
The Eurostat Crime and Criminal Justice dataset provides yearly statistics on the citizens and non-citizens within the justice system across European Union member states, covering the period from 2008 to 2022. It includes data on suspicion and conviction rates per 1,000 inhabitants, offering insights into both the number of individuals suspected of crimes and those convicted. By distinguishing between citizens and non-citizens, this dataset sheds light on potential disparities in how these groups are treated within the legal system. Such information is crucial for understanding systemic inequities and evaluating the impact of policies on different demographics. 

Manifesto Project Dataset
The Manifesto Project dataset offers a systematic analysis of political party manifestos across various countries, including EU member states. Spanning elections from 1946 to 2017 (with country-specific coverage), it captures the percentage of text devoted to key themes such as "law and order," "national security," and "national values." This dataset is particularly valuable for studying the evolution of political discourse over time and across contexts. The dataset’s coding of text into quantifiable measures makes it a powerful tool for understanding the role of party platforms in shaping public opinion and influencing policy. Its detailed historical scope enables longitudinal studies of political ideologies and their relationship with contemporary governance trends.

EU Political Barometer Dataset
The EU Political Barometer dataset provides bi-monthly data on public opinion and political preferences across EU member states from 2019 to 2023. It tracks ideological shifts, political attitudes, and public reactions to major societal events and political campaigns. Key indicators include changes in support for various ideologies and parties, offering a granular view of how public sentiment evolves over time. This dataset is particularly useful for analyzing short-term trends and understanding the relationship between political discourse and public opinion. By examining fluctuations in attitudes during specific events or election campaigns, we can identify patterns in voter behavior and ideological alignment. Its frequent updates make it a critical resource for real-time political analysis and policy evaluation.

5- Description table

```{r}


# Function to calculate dataset summary
get_summary <- function(data, name, year_column, date_format, country_column) {
  # Normalize the year format
  years <- tryCatch({
    if (date_format == "year") {
      data[[year_column]]
    } else if (date_format == "year-month-day") {
      year(as.Date(data[[year_column]]))
    } else if (date_format == "day/month/year") {
      year(dmy(data[[year_column]]))
    } else {
      NA
    }
  }, error = function(e) NA)
  
  # Check validity of years and calculate range
  if (all(is.na(years))) {
    years_covered <- "Invalid Year Column"
  } else {
    years_covered <- paste(range(years, na.rm = TRUE), collapse = " - ")
  }
  
  # Count unique countries
  num_countries <- if (country_column %in% names(data)) {
    length(unique(data[[country_column]]))
  } else {
    0
  }
  
  # Compute summary statistics
  tibble(
    Dataset = name,
    Rows = nrow(data),
    Columns = ncol(data),
    `Years Covered` = years_covered,
    `Number of Countries` = num_countries,
    `Total Datapoints` = nrow(data) * ncol(data)
  )
}

# Example datasets and their year and country columns
datasets <- list(
  list(name = "Crime Dataset", data = crime, year_column = "TIME_PERIOD", date_format = "year", country_column = "geo"),
  list(name = "Ideology Dataset", data = ideology, year_column = "date", date_format = "year-month-day", country_column = "country_name"),
  list(name = "Manifesto Dataset", data = manifesto, year_column = "edate", date_format = "day/month/year", country_column = "countryname")
)

# Generate summaries
summaries <- lapply(datasets, function(ds) {
  get_summary(ds$data, ds$name, ds$year_column, ds$date_format, ds$country_column)
})

# Combine all summaries into one table
summary_table <- bind_rows(summaries)

# Display the summary table
print(summary_table)
```

