---
title: "Crime, Ideology, and Manifestos: A Comparative Study of European Political Trends"
author: "Alex Martinez & Serry Ezbidi"
format:
  html:
    number-sections: true
  pdf:
    number-sections: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE 
)
```

```{r, include=FALSE}
here::i_am("Data-Management-Project.Rproj")
library(here)
library(vroom)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(eurostat)
library(lubridate)
library(knitr)
library(tibble)
library(kableExtra)
library(countrycode)
library(tinytex)
library(gridExtra)
library(gridExtra)
library(grid)
```

```{r, include=FALSE}
crime <- get_eurostat("crim_just_ctz", time_format = "num", stringsAsFactors = TRUE)
manifesto <- vroom(("https://manifesto-project.wzb.eu/down/data/2024a/datasets/MPDataset_MPDS2024a.csv"))
ideology <- vroom(("https://eupoliticalbarometer.uc3m.es/api/ideologyDownload"))
pop <- get_eurostat("migr_pop2ctz", time_format = "num", stringsAsFactors = TRUE)
```

# **Introduction**

## *Link to GitHub*

<a href="https://github.com/alexmar313/Data-Management-Project.git" target="_blank">GitHub Repository</a>

## *Sources download links*

- <a href="https://ec.europa.eu/eurostat/databrowser/view/crim_just_ctz/default/table?lang=en" target="_blank">Eurostat Crime Dataset</a>
- <a href="https://ec.europa.eu/eurostat/databrowser/view/migr_pop2ctz/default/table?lang=en&category=demo.demo_pop">Eurostat Population Dataset</a>
- <a href="https://manifesto-project.wzb.eu/down/data/2024a/datasets/MPDataset_MPDS2024a.csv" target="_blank">Manifesto Dataset</a>
- <a href="https://eupoliticalbarometer.uc3m.es/api/ideologyDownload" target="_blank">Ideology Dataset</a>

## *Websites links from which data was extracted*

- <a href="https://ec.europa.eu/eurostat/databrowser/view/crim_just_ctz/default/table?lang=en" target="_blank">Eurostat Crime Dataset</a>
- <a href="https://ec.europa.eu/eurostat/api/dissemination/sdmx/2.1/data/migr_pop2ctz?format=TSV&compressed=true">Eurostat Population Dataset<a>
- <a href="https://manifesto-project.wzb.eu/datasets" target="_blank">Manifesto Project Dataset</a>
- <a href="https://eupoliticalbarometer.uc3m.es/dashboard/ideology" target="_blank">EU Political Barometer Dashboard</a>

# **Data Description**

## *Eurostat Crime and Criminal Justice Dataset*
The Eurostat Crime and Criminal Justice dataset provides yearly statistics on the citizens and non-citizens within the justice system across European Union member states, covering the period from 2008 to 2022. It includes data on suspicion and conviction rates per 1,000 inhabitants, offering insights into both the number of individuals suspected of crimes and those convicted. By distinguishing between citizens and non-citizens, this dataset sheds light on potential disparities in how these groups are treated within the legal system. Such information is crucial for understanding systemic inequities and evaluating the impact of policies on different demographics. 

## *Eurostat Population*
The Eurostat Migration dataset provides annual data on the population of citizens and non-citizens across EU member states from 2008 to 2022. Non-citizens include foreigners, stateless individuals, and those with unknown status. This dataset allows for detailed analysis of demographic compositions and is crucial for understanding disparities between citizens and non-citizens. We aim to use it to determine the number of citizens and non-citizens in each country, highlighting migration trends and their societal implications.

## *Manifesto Project Dataset*
The Manifesto Project dataset offers a systematic analysis of political party manifestos across various countries, including EU member states. Spanning elections from 1946 to 2017 (with country-specific coverage), it captures the percentage of text devoted to key themes such as "law and order," "national security," and "national values." This dataset is particularly valuable for studying the evolution of political discourse over time and across contexts. The dataset’s coding of text into quantifiable measures makes it a powerful tool for understanding the role of party platforms in shaping public opinion and influencing policy. Its detailed historical scope enables longitudinal studies of political ideologies and their relationship with contemporary governance trends.

## *EU Political Barometer Dataset*
The EU Political Barometer dataset provides bi-monthly data on public opinion and political preferences across EU member states from 2019 to 2023. It tracks ideological shifts, political attitudes, and public reactions to major societal events and political campaigns. Key indicators include changes in support for various ideologies and parties, offering a granular view of how public sentiment evolves over time. This dataset is particularly useful for analyzing short-term trends and understanding the relationship between political discourse and public opinion. By examining fluctuations in attitudes during specific events or election campaigns, we can identify patterns in voter behavior and ideological alignment. Its frequent updates make it a critical resource for real-time political analysis and policy evaluation.

## *Description Table*

```{r, echo=FALSE}
get_summary <- function(data, name, year_column, date_format, country_column) {

  years <- tryCatch({
    if (date_format == "year") {
      data[[year_column]]
    } else if (date_format == "year-month-day") {
      year(as.Date(data[[year_column]]))
    } else if (date_format == "day/month/year") {
      year(dmy(data[[year_column]]))
    } else {
      NA
    }
  }, error = function(e) NA)

  if (all(is.na(years))) {
    years_covered <- "Invalid Year Column"
  } else {
    years_covered <- paste(range(years, na.rm = TRUE), collapse = " - ")
  }
  
  num_countries <- if (country_column %in% names(data)) {
    length(unique(data[[country_column]]))
  } else {
    0
  }
  
  tibble(
    Dataset = name,
    Rows = nrow(data),
    Columns = ncol(data),
    `Years Covered` = years_covered,
    `Number of Countries` = num_countries,
    `Total Datapoints` = nrow(data) * ncol(data)
  )
}

datasets <- list(
  list(name = "Crime Dataset", data = crime, year_column = "TIME_PERIOD", date_format = "year", country_column = "geo"),
  list(name = "Ideology Dataset", data = ideology, year_column = "date", date_format = "year-month-day", country_column = "country_name"),
  list(name = "Manifesto Dataset", data = manifesto, year_column = "edate", date_format = "year-month-day", country_column = "countryname"),
  list(name = "Population Dataset", data = pop, year_column = "TIME_PERIOD", date_format = "year", country_column = "geo")
)

summaries <- lapply(datasets, function(ds) {
  get_summary(ds$data, ds$name, ds$year_column, ds$date_format, ds$country_column)
})

summary_table <- bind_rows(summaries) 
knitr::kable(summary_table)

```

# **Description of Variables of Interest**

## *Crime Dataset*

- *"leg_stat"*: "PER_SUSP" indicates individuals who are suspected of committing crimes, "PER_CNV" indicates individuals who are convicted of crimes.

- *"citizen"*: "NAT" indicates nationals (citizens of the reporting country), "FOR" represents foreigners (non-citizens).

## *Population Dataset*

- *"citizen"*: "FOR_STLS" indicates individuals who are foreigners or stateless, "NAT" indicates individuals who have a citizenship.

- *"sex"*: "F" & "M" indicates whether individuals are women or men, "T" represents the total.

## *Ideology Dataset*

- *"left_ideology"*: numeric score (0-10) representing the left-wing ideological positioning in the country, where a higher value corresponds to stronger left ideology.

- *"right_ideology"*: numeric score (0-10) representing the right-wing ideological positioning in the country, where a higher value corresponds to stronger right ideology.

- *"dalton"*: named after the political scientist Russell Dalton, a numeric score (0-10) that shows the degree of ideological polarization in a country, where a higher score corresponds to higher polarization.

## *Manifesto Dataset*

- *"per101 to per109"*: represent the percentage of the political party's manifesto dedicated to specific themes related to national security, crime, and immigration. They focus on topics like law and order, national security, crime prevention, and the role of the state in dealing with security threats. Specifically: per101: Law and Order, per102: National Security, per103: Crime and Punishment, per104: Prison and Penal System, per105: Immigration, per106: International Relations (related to security), per107: Economic Issues, per108: Welfare and Social Issues, per109: Cultural and National Identity

- *"per201 to per204"*: focus on economic policies, social support, and public services, potentially linking to discussions about immigration’s impact on the economy and social welfare. Specifically, per201: Economic Growth, per202: Employment, per203: Social Security, per204: Public Services

- *"per301 to per305"*: focus on social welfare, social issues, and public goods, which might also intersect with debates around immigration and crime in relation to societal well being and state responsibility. Specifically, per301: Social Welfare, per302: Education, per303: Health Care, per304: Family Support, per305: Environment and Sustainability.

# **Data Cleaning Process**

## *Cleaning of Eurostat Data, Crime*
```{r, include=FALSE}
crime <- crime %>%
  filter(citizen != "TOTAL" & leg_stat != "PER_PRSC" & unit != "P_HTHAB")
```
- *Filtering and cleaning:* remove irrelevant categories (e.g., "TOTAL" citizens and "PER_PRSC" status), keeping only the data as total number of citizens and non-citizens suspicted and convicted to crimes.

```{r, include=FALSE}
crime <- crime %>%
  mutate(geo = countrycode(geo, origin = "iso2c", destination = "country.name"))
crime <- crime %>%
  filter(!is.na(geo))
```
- *Country name assignment:* convert country codes to full country names using the "countrycode" package.

```{r, include=FALSE}
crime <- crime %>%
  mutate(Category = case_when(
    citizen == "NAT" & leg_stat == "PER_CNV" ~ "Convicted Citizens",
    citizen == "FOR" & leg_stat == "PER_CNV" ~ "Convicted Non-Citizens",
    citizen == "NAT" & leg_stat == "PER_SUSP" ~ "Suspected Citizens",
    citizen == "FOR" & leg_stat == "PER_SUSP" ~ "Suspected Non-Citizens",
    TRUE ~ NA_character_  
  ))
```
- *Category creation:* classify data into four categories based on citizenship and legal status: "Convicted Citizens," "Convicted Non-Citizens," "Suspected Citizens," and "Suspected Non-Citizens."

```{r, include=FALSE}
crime <- crime %>%
  select(geo, TIME_PERIOD, Category, values)
crime <- crime %>%
  rename(
    year = TIME_PERIOD,
    country = geo
  )
```
- *Drop irrelevant variables:* rename and include only variables of interest: country, date, citizenship & legal status category, and crime rates.

```{r, include=FALSE}
crime <- crime %>%
  pivot_wider(
    names_from = Category,
    values_from = values    
  )
```
- *Data reshaping:* pivot the data to have a single row per country per year, with columns for each of the four categories (convicted and suspected citizens/non-citizens).


## *Cleaning of Eurostat Data, Population*
```{r, include=FALSE}
Tpop <- pop %>%
  filter(age == "TOTAL", sex == "T", citizen != "UNK", citizen != "TOTAL") %>%
  rename(year = TIME_PERIOD) 
```
- *Filtering and cleaning:* remove irrelevant category "UNK" or "TOTAL" citizens), keeping only the data as total numbers, not separating per sex or age.

```{r, include=FALSE}
Tpop <- Tpop %>%
  mutate(country = countrycode(geo, origin = "iso2c", destination = "country.name", warn = FALSE)) %>%
  filter(!is.na(country)) 
```
- *Country name assignment:* convert country codes to full country names using the "countrycode" package.

```{r, include=FALSE}
Tpop <- Tpop %>%
  group_by(country, year, citizen) %>%
  summarise(
    population = sum(values, na.rm = TRUE),
    .groups = "drop"
  )
```
- *Summarize the data:* group the data by country, year, and citizen and calculate the total population for each group.

```{r, include=FALSE}
Tpop <- Tpop %>%
  pivot_wider(
    names_from = citizen,
    values_from = population,
    values_fill = 0
  ) %>%
  mutate(
    population_citizens = NAT,
    population_non_citizens = FOR_STLS,
    population_total = NAT + FOR_STLS
  )
```
-*Reshape Data for Citizens and Non-Citizens:* Reshapes the data to separate populations of citizens (NAT) and non-citizens (FOR_STLS)


## *Cleaning of Manifesto Data, manifesto*
```{r, include=FALSE}
manifesto <- manifesto %>%
  mutate(edate = dmy(edate)) %>%
  filter(edate >= as.Date("2008-01-01"))
```
- *Date filtering:* remove data before 2008 (the first available year in the Crime dataset).

```{r, include=FALSE}
manifesto <- manifesto %>%
  select(countryname, partyname, edate, per101, per102, per103, per104, per105, per106, per107, per108, per109, per201, per203, per202, per204, per301, per302, per303, per304, per305)
```
- *Column selection:* Keeps only relevant columns in the manifesto dataset (e.g., political themes related to law and order, national security, immigration).

```{r, include=FALSE}
manifesto <- manifesto %>%
  mutate(date = year(edate)) 

manifesto <- manifesto %>%
  select(countryname, date, per101, per102, per103, per104, per105, per106, per107, per108, per109, per201, per203, per202, per204, per301, per302, per303, per304, per305)

manifesto <- manifesto %>%
  group_by(countryname, date) %>%
  summarise(
    per101 = mean(per101, na.rm = TRUE),
    per102 = mean(per102, na.rm = TRUE),
    per103 = mean(per103, na.rm = TRUE),
    per104 = mean(per104, na.rm = TRUE),
    per105 = mean(per105, na.rm = TRUE),
    per106 = mean(per106, na.rm = TRUE),
    per107 = mean(per107, na.rm = TRUE),
    per108 = mean(per108, na.rm = TRUE),
    per109 = mean(per109, na.rm = TRUE),
    per201 = mean(per201, na.rm = TRUE),
    per202 = mean(per202, na.rm = TRUE),
    per203 = mean(per203, na.rm = TRUE),
    per204 = mean(per204, na.rm = TRUE),
    per301 = mean(per301, na.rm = TRUE),
    per302 = mean(per302, na.rm = TRUE),
    per303 = mean(per303, na.rm = TRUE),
    per304 = mean(per304, na.rm = TRUE),
    per305 = mean(per305, na.rm = TRUE)
  ) %>%
  ungroup()

manifesto <- manifesto %>%
  rename(year = date,
         country = countryname)
```
- *Group by date and country:* extract the year from the date and group the data by country and year to compute the average percentage of the manifesto dedicated to specific themes like law and order, national security, and immigration.


## *Cleaning of Barometer Data, ideology*
```{r, include=FALSE}
ideology <- ideology %>%
  filter(data_type != "ewma")
```
- *Filter invalid data:* removes data related to "ewma" (exponentially weighted moving average) and retain only real ideological scores.

```{r, include=FALSE}
ideology <- ideology %>%
  filter(date < as.Date("2023-01-01"))
```
- *Date filtering:* filter the dataset to include only data up until 2023.

```{r, include=FALSE}
ideology <- ideology %>%
  mutate(year = year(date))

ideology <- ideology %>%
  group_by(country_name, year) %>%
  summarise(
    left_ideology = mean(left_ideology, na.rm = TRUE),
    right_ideology = mean(right_ideology, na.rm = TRUE),
    dalton = mean(dalton, na.rm = TRUE)
  ) %>%
  ungroup()

ideology <- ideology %>%
  rename(country = country_name)
```
_ *Ideology Data:* extract the year from the date and groups the data by country and year to calculate the average ideological scores (left, right, and Dalton's polarization index).


## *Joint Cleaning*
```{r, include=FALSE}
crime <- crime %>%
  semi_join(Tpop, by = c("country" = "country"))
Tpop <- Tpop %>%
  semi_join(crime, by = c("country" = "country"))
ideology <- ideology %>%
  semi_join(crime, by = c("country" = "country"))
manifesto <- manifesto %>%
  semi_join(crime, by = c("country" = "country"))
```
- *Data intersection:* ensure that only countries present in the Crime and Population are kept in the Manifesto and Ideology data sets. This is achieved through a semi_join to keep only the common countries.


## *Merge data sets*
```{r, include=FALSE}
crime$year <- as.numeric(crime$year)
manifesto$year <- as.numeric(manifesto$year)
ideology$year <- as.numeric(ideology$year)

merged_data <- full_join(crime, manifesto, by = c("country", "year")) %>%
  full_join(ideology, by = c("country", "year"))
```
-*Merge Crime, Manifesto and Ideology*: previously renamed columns in all three data sets for consistency and merge them into one unified data set. "full_join" is used to ensure no data is lost while combining them based on country and year.

```{r, include=FALSE}
merged_data <- Tpop %>%
  inner_join(merged_data, by = c("country", "year"))
merged_data <- merged_data %>%
  mutate(
    convicted_p_c = (`Convicted Citizens` / `population_citizens`) * 100,
    convicted_p_nc = (`Convicted Non-Citizens` / `population_non_citizens`) * 100,
    suspected_p_c = (`Suspected Citizens` / `population_citizens`) * 100,
    suspected_p_nc = (`Suspected Non-Citizens` / `population_non_citizens`) * 100
  )
```
-*Merge also with Population Data:* add the fourth data set. It would be informative to present the data as percentages in addition to raw numbers. For example, we can display the proportion of suspected non-citizens relative to the total population of non-citizens, and compare this to the proportion of suspected citizens relative to the total population of citizens. We hence also calculate percentages for convictions and suspicions for both citizens and non-citizens


# **Description of the Research Question**

## *Research question:* What is the association between immigration and crime suspicions and convictions in the EU, considering political ideology?

The research explores the relationship between immigration status and crime outcomes (suspicion and conviction rates) in EU countries. It questions whether claims by political entities, particularly right-wing parties, that immigration drives up crime rates hold true, and explores how other intervening factors, such as the demographic profile of immigrants, might contribute to this perception.

Background context and relevance
Right-wing political narratives often link immigration with rising crime rates without acknowledging broader contextual factors. For example, statistical evidence suggests that most crimes are committed by young males—a group overrepresented among immigrant populations. Across the EU, a significant proportion of non-citizens are young males: non-national men aged 20–49 constitute 29% of their demographic group, compared to 18% for nationals. Additionally, 54–60% of unauthorized immigrants are male, with a majority under 35 years old. Such demographic realities could skew perceptions of immigrant involvement in crime when not carefully controlled for.
*Source: <a href="https://ec.europa.eu/eurostat/web/interactive-publications/migration-2023" target="_blank">Migration 2023</a>

This research aims to integrate crime statistics and demographic data with measures of political ideology to disentangle these associations. By doing so, it challenges simplistic narratives and aims to explore whether shifts in political rhetoric influence crime outcomes or merely exacerbate perceptions of immigrant criminality.

## *Data and methodology*
1. Crime Data:
   Eurostat provides comprehensive data on annual suspicion and conviction rates per 1,000 residents, dis-aggregated by citizenship (non-citizens vs. citizens) for 2008–2022. This allows for trend analysis of disparities between groups.

2. Political ideology data:
   - Manifesto Project Database: Tracks the content of political party platforms during election years, focusing on themes such as “law and order” and “security.”  
   - EU Political Barometer: Captures public opinion on political ideologies monthly from 2019–2023.

A Difference-in-Differences (DiD) approach will compare changes in crime-related outcomes before and after election periods with shifts in ideological rhetoric. Non-election years serve as a control group, isolating the causal effect of political discourse.

## *Core issues addressed* 
The research will investigate:
1. Whether increased suspicion of immigrant crime correlates with heightened right-wing rhetoric.
2. If young male over representation among immigrants, rather than immigrant status itself, explains these crime rates.  
3. Do conviction rates—representing judicial outcomes—reflect ideological trends, or are they more stable and less subject to political influence?

## *Significance and expected outcomes*
This study seeks to clarify how political discourse shapes public perceptions and crime outcomes involving immigrants. Preliminary hypotheses suggest that while suspicion rates rise in line with ideological shifts, conviction rates remain relatively stable, suggesting biases in suspicion rather than evidence-based outcomes. By emphasizing intervening factors like demographics, the research highlights the importance of nuanced policymaking and avoids perpetuating harmful stereotypes.

# **Data Visualisation**

To address the research question of this project, the analysis is structured to explore the relationship between immigration and crime, as well as the broader sociopolitical factors that may influence this dynamic. The analysis begins by visually examining the primary variables of interest: the target variable, crime, and the main explanatory variable, immigration. For each of these, we decompose trends over the years 2008 to 2022. This step allows us to identify patterns and changes over time, providing an essential context for interpreting the relationship between immigration and crime. By focusing on visual representation, we aim to make the data more intuitive and accessible, setting the foundation for subsequent statistical analysis.

The next stage involves running regression analyses to quantitatively investigate the association between changes in the percentage of non-citizens in a country and the rates of conviction and suspicion of crime. This step is critical to isolating and understanding the potential link between immigration and crime, while controlling for other relevant factors. These regressions provide robust statistical insights into the strength and direction of the relationship, allowing us to determine whether increases in immigration are associated with significant changes in crime-related outcomes.

Beyond the primary relationship between immigration and crime, we expand the analysis to include additional variables that may shape or mediate this relationship. Two key factors considered are the level of political polarization within a country, captured by the Dalton index, and the prevalence of right-wing ideology. Political polarization reflects the extent to which opinions are divided within a society, which could influence perceptions of immigration and its relationship to crime. Similarly, the prevalence of right-wing ideology may shape public discourse and policy responses to immigration, potentially affecting crime rates or how they are reported. Incorporating these variables adds depth to our analysis, enabling us to explore how political and ideological contexts intersect with immigration and crime.

Finally, we analyze the impact of immigration-related terminology in political party manifestos during election years in the countries included in our sample. By studying the language used in these documents, we aim to capture the role of political rhetoric in shaping public perceptions and potentially influencing crime-related outcomes. This aspect of the analysis allows us to connect trends in immigration and crime to broader sociopolitical narratives, offering a more holistic understanding of the issue.

Through this multi-faceted approach, which combines visual analysis, regression models, and contextual exploration of political and ideological factors, the analysis seeks to uncover nuanced insights into the relationship between immigration and crime. By situating this relationship within its broader societal and political context, we provide a richer and more comprehensive framework for addressing the research question.

When shifting to this relative measure, we see that non-citizens are shown to have higher conviction and suspicion rates compared to citizens. Specifically, the average percentage of non-citizens suspected of crimes in our sample is around 4%. This trend includes a notable spike in 2014, where suspicion rates sharply increase, before returning to approximately 4% in subsequent years. In contrast, the suspicion rate for citizens remains much lower and more stable, averaging between 0-1%, although a slight upward spike is observed in 2013.

Similarly, conviction rates exhibit distinct patterns for citizens and non-citizens. For non-citizens, the average percentage convicted of crimes is around 3% between 2008 and 2014. This figure declines steadily in later years, reaching approximately 1.5% by 2020. For citizens, the average conviction rate is initially around 1% during the same period but gradually decreases, reaching almost zero by 2022. These trends suggest a pronounced difference in the treatment or behavior of citizens versus non-citizens within the criminal justice system.


## *Visualization of the Target Variable - Crime*
### *Total Crime Suspicion Rates Over Time for Citizens and Non-Citizens in Selected Countries*
```{r, echo=FALSE}
filtered_data <- crime %>%
  filter(country %in% c("France", "Germany", "Austria", "Bulgaria"))

ggplot(filtered_data, aes(x = year)) +
  geom_line(aes(y = `Suspected Non-Citizens`, color = "Suspected Non-Citizens"), linewidth = 1) +
  geom_line(aes(y = `Suspected Citizens`, color = "Suspected Citizens"), linewidth = 1) +
  facet_wrap(~country, scales = "free_y") + 
  labs(
    title = "Crime Suspicion Over Time",
    x = "Year",
    y = "Number of People",
    color = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"  # Move the legend below the graph
  ) +
  scale_color_manual(
    values = c(
      "Suspected Non-Citizens" = "red", 
      "Suspected Citizens" = "orange"
    )
  )
```

### *Total Crime Conviction Rates Over Time for Citizens and Non-Citizens in Selected Countries*
```{r, echo=FALSE}
ggplot(filtered_data, aes(x = year)) +
  geom_line(aes(y = `Convicted Non-Citizens`, color = "Convicted Non-Citizens"), linewidth = 1) +
  geom_line(aes(y = `Convicted Citizens`, color = "Convicted Citizens"), linewidth = 1) +
  facet_wrap(~country, scales = "free_y") +  # Split by country
  labs(
    title = "Crime Conviction Over Time",
    x = "Year",
    y = "Number of People",
    color = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"  # Move the legend below the graph
  ) +
  scale_color_manual(
    values = c(
      "Convicted Non-Citizens" = "red", 
      "Convicted Citizens" = "orange"
    )
  )
```

### *Average Crime Suspicion Rates Over Time for Citizens and Non-Citizens as a Percentage of their Population Share*
```{r, echo=FALSE}
merged_data %>%
  group_by(year) %>%
  summarise(
    avg_citizens = mean(suspected_p_c, na.rm = TRUE),
    avg_non_citizens = mean(suspected_p_nc, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = c(avg_citizens, avg_non_citizens),
    names_to = "group",
    values_to = "value_percentage"
  ) %>%
  mutate(group = ifelse(group == "avg_citizens", "Citizens", "Non-Citizens")) %>%
  ggplot(aes(x = year, y = value_percentage, color = group, linetype = group)) +
  geom_line(size = 1.2) +
  labs(
    title = "Average Percentage of Suspected Citizens and Non-Citizens",
    x = "Year",
    y = "Average Percentage",
    color = "Group",
    linetype = "Group"
  ) +
  theme_minimal()

```

### *Average Crime Conviction Rates Over Time for Citizens and Non-Citizens as a Percentage of their Population Share*
```{r, echo=FALSE}
merged_data %>%
  group_by(year) %>%
  summarise(
    avg_citizens = mean(convicted_p_c, na.rm = TRUE),
    avg_non_citizens = mean(convicted_p_nc, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = c(avg_citizens, avg_non_citizens),
    names_to = "group",
    values_to = "value_percentage"
  ) %>%
  mutate(group = ifelse(group == "avg_citizens", "Citizens", "Non-Citizens")) %>%
  ggplot(aes(x = year, y = value_percentage, color = group, linetype = group)) +
  geom_line(size = 1.2) +
  labs(
    title = "Average Percentage of Convicted Citizens and Non-Citizens",
    x = "Year",
    y = "Average Percentage",
    color = "Group",
    linetype = "Group"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
countries <- c("France", "Germany", "Austria", "Bulgaria")

for(country in countries) {

  p1 <- merged_data %>%
    filter(country == country) %>%
    group_by(year) %>%
    summarise(
      convicted = mean(convicted_p_c, na.rm = TRUE),
      suspected = mean(suspected_p_c, na.rm = TRUE)
    ) %>%
    pivot_longer(
      cols = c(convicted, suspected),
      names_to = "type",
      values_to = "percentage"
    ) %>%
    mutate(type = tools::toTitleCase(type)) %>%
    ggplot(aes(x = year, y = percentage, color = type, linetype = type)) +
    geom_line(linewidth = 1.2) +
    labs(
      title = paste(country, "- Citizens"),
      x = "Year",
      y = "Percentage",
      color = "Type",
      linetype = "Type"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  p2 <- merged_data %>%
    filter(country == country) %>%
    group_by(year) %>%
    summarise(
      convicted = mean(convicted_p_nc, na.rm = TRUE),
      suspected = mean(suspected_p_nc, na.rm = TRUE)
    ) %>%
    pivot_longer(
      cols = c(convicted, suspected),
      names_to = "type",
      values_to = "percentage"
    ) %>%
    mutate(type = tools::toTitleCase(type)) %>%
    ggplot(aes(x = year, y = percentage, color = type, linetype = type)) +
    geom_line(linewidth = 1.2) +
    labs(
      title = paste(country, "- Non-Citizens"),
      x = "Year",
      y = "Percentage",
      color = "Type",
      linetype = "Type"
    ) +
    theme_minimal() + 
    theme(legend.position = "bottom")

  grid.arrange(p1, p2, ncol = 2, top = "Convicted vs Suspected Percentages")
}

```

The graphs reveal several important trends when examining the conviction and suspicion rates of citizens and non-citizens over time. First, we observe that in absolute terms, the number of citizens convicted or suspected of crimes is consistently higher than that of non-citizens. This finding holds true across our subset of the sample, which includes . However, this observation alone does not provide the full picture, as it does not account for the relative size of the citizen and non-citizen populations in each country. To better address our research question, we instead examine the number of convicted and suspected individuals as a percentage of the total population of citizens or non-citizens within a country.

<<<<<<< HEAD
# **Data Analysis**

## **Relationship between Percentage of Migrants and Total Convictions**
```{r, echo=FALSE}
merged_data_final <- merged_data %>%
  mutate(
    total_convicted = `Convicted Citizens` + `Convicted Non-Citizens`,
    total_suspected = `Suspected Citizens` + `Suspected Non-Citizens`,
    percentage_nc_pop = (population_non_citizens / (population_non_citizens + population_citizens)) * 100
  )

convicted_analysis_data <- merged_data_final %>%
  select(country, year, percentage_nc_pop, total_convicted)

ggplot(convicted_analysis_data, aes(x = percentage_nc_pop, y = total_convicted, color = country)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "black", linetype = "dashed") +
  labs(
    title = "Relationship between the Percentage of Non-Citizens and Total Conviction",
    x = "Percentage of Non-Citizens in Total Population",
    y = "Total Convictions"
  ) +
  theme_minimal()

model_convicted <- lm(total_convicted ~ percentage_nc_pop, 
                     data = merged_data_final)

Conviction_Model <- data.frame(
  Term = c("Intercept", "% Non-Citizens"),
  Estimate = coef(model_convicted),
  `Std.Error` = summary(model_convicted)$coefficients[,2],
  `t.value` = summary(model_convicted)$coefficients[,3],
  `p.value` = summary(model_convicted)$coefficients[,4]
) %>%
  kbl(caption = "Regression Results: Total Convicted Persons",
      digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

Conviction_Model

model_convicted <- lm(total_convicted ~ percentage_nc_pop, data = merged_data_final)
summary(model_convicted)
```

## **Relationship between Percentage of Migrants and Total Suspicions**
```{r, echo=FALSE}
suspected_analysis_data <- merged_data_final %>%
  select(country, year, percentage_nc_pop, total_suspected)

ggplot(suspected_analysis_data, aes(x = percentage_nc_pop, y = total_suspected, color = country)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "black", linetype = "dashed") +
  labs(
    title = "Relationship between the Percentage of Non-Citizens and Total Suspicions",
    x = "Percentage of Non-Citizens in Total Population",
    y = "Total Suspicions"
  ) +
  theme_minimal()

model_suspected <- lm(total_suspected ~ percentage_nc_pop, 
                     data = merged_data_final)

Supspicion_Model <- data.frame(
  Term = c("Intercept", "% Non-Citizens"),
  Estimate = coef(model_suspected),
  `Std.Error` = summary(model_suspected)$coefficients[,2],
  `t.value` = summary(model_suspected)$coefficients[,3],
  `p.value` = summary(model_suspected)$coefficients[,4]
) %>%
  kbl(caption = "Regression Results: Total Suspected Persons",
      digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

Supspicion_Model

model_suspected <- lm(total_suspected ~ percentage_nc_pop, data = merged_data_final)
summary(model_suspected)
```


## **Relationship between Percentage of Migrants and Suspicions of Non-citizens**
- Removed Bulgaria as it was a clear outlier
```{r, echo=FALSE}
suspected_analysis_data <- merged_data_final %>%
  filter(country != "Bulgaria") %>% 
  select(country, year, percentage_nc_pop, suspected_p_nc)

ggplot(suspected_analysis_data, aes(x = percentage_nc_pop, y = suspected_p_nc, color = country)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "black", linetype = "dashed") +
  labs(
    title = "Relationship between the Percentage of Migrants and Suspicions of Non-citizens",
    x = "Percentage of Non-Citizens in Total Population",
    y = "Percentage of Non-citizens Suspected"
  ) +
  theme_minimal()

model_suspected_nc <- lm(suspected_p_nc ~ percentage_nc_pop, 
                     data = merged_data_final  %>% filter(country != "Bulgaria"))

Supspicion_Model_nc <- data.frame(
  Term = c("Intercept", "% Non-Citizens"),
  Estimate = coef(model_suspected),
  `Std.Error` = summary(model_suspected)$coefficients[,2],
  `t.value` = summary(model_suspected)$coefficients[,3],
  `p.value` = summary(model_suspected)$coefficients[,4]
) %>%
  kbl(caption = "Regression Results: Total Suspected Persons",
      digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

Supspicion_Model_nc
```

## **Relationship between Percentage of Migrants and Convictions of Non-citizens**
```{r, echo=FALSE}
suspected_analysis_data <- merged_data_final %>%
  select(country, year, percentage_nc_pop, convicted_p_nc)

ggplot(suspected_analysis_data, aes(x = percentage_nc_pop, y = convicted_p_nc, color = country)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "black", linetype = "dashed") +
  labs(
    title = "Relationship between the Percentage of Migrants and Percentage Migrants Convictions",
    x = "Percentage of Non-Citizens in Total Population",
    y = "Percentage of Non-citizens Convicted"
  ) +
  theme_minimal()

model_convicted_nc <- lm(convicted_p_nc ~ percentage_nc_pop, 
                     data = merged_data_final)

Conviction_Model_nc <- data.frame(
  Term = c("Intercept", "% Non-Citizens"),
  Estimate = coef(model_suspected),
  `Std.Error` = summary(model_suspected)$coefficients[, 2],
  `t.value` = summary(model_suspected)$coefficients[, 3],
  `p.value` = summary(model_suspected)$coefficients[, 4]
) %>%
  kbl(
    caption = "Regression Results: Total Suspected Persons",
    digits = 3,
    align = 'c'
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)

Conviction_Model_nc
```

## **Relationship between Polarization and Total Convictions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = dalton, y = total_convicted)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Dalton and Total Convicted", 
       x = "Dalton Index", 
       y = "Total Convicted") +
  theme_minimal()

model_with_polynomial <- lm(total_convicted ~ dalton + poly(year, 2), data = merged_data_final)
summary(model_with_polynomial)
```

## **Relationship between Polarization and Percentage Migrants Convictions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = dalton, y = convicted_p_nc)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Dalton and Percentage Migrants Conviction", 
       x = "Dalton Index", 
       y = "Percentage Convicted Migrants") +
  theme_minimal() +
   ylim(0, 5)

model_with_polynomial <- lm(convicted_p_nc ~ dalton + poly(year, 2), data = merged_data_final)
summary(model_with_polynomial)
```

## **Relationship between Polarization and Percentage Migrants Suspicions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = dalton, y = suspected_p_nc)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Dalton and Percentage Migrants Suspicions", 
       x = "Dalton Index", 
       y = "Percentage Suspected Migrants") +
  theme_minimal() +
  ylim(0, 10)

model_with_polynomial <- lm(suspected_p_nc ~ dalton + poly(year, 2), data = merged_data_final)
summary(model_with_polynomial)
```


## **Relationship between Right Ideology and Total Convictions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = right_ideology, y = total_convicted)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Right Ideology and Total Conviction", 
       x = "Right Ideology", 
       y = "Total Convictions") +
  theme_minimal()

model_with_ideology <- lm(total_convicted ~ right_ideology + year, data = merged_data_final)
summary(model_with_ideology)

```

## **Relationship between Right Ideology and Percentage Migrants Convictions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = right_ideology, y = convicted_p_nc)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Right Ideology and Percentage of Migrants Convicted", 
       x = "Right Ideology", 
       y = "Non-citiznes Convictions") +
  theme_minimal() +
  ylim(0, 5)

model_with_ideology <- lm(convicted_p_nc ~ right_ideology + year, data = merged_data_final)
summary(model_with_ideology)

```


## **Relationship between Right Ideology and Percentage Migrants Suspicions**
```{r, echo=FALSE}
ggplot(merged_data_final, aes(x = right_ideology, y = suspected_p_nc)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship Between Right Ideology and Percentage of Migrants Suspected", 
       x = "Right Ideology", 
       y = "Non-citiznes Suspicons") +
  theme_minimal() +
  ylim(0, 5)

model_with_ideology <- lm(suspected_p_nc ~ right_ideology + year, data = merged_data_final)
summary(model_with_ideology)
```

```{r, echo=FALSE}
merged_data_for_immigration_visual <- merged_data %>%
  filter(country %in% c("France", "Germany", "Austria", "Bulgaria"))

ggplot(merged_data_for_immigration_visual, aes(x = year)) +
  geom_line(aes(y = (FOR_STLS/population_total)*100, color = "Percentage of Non-Citizens"), linewidth = 1) +
  facet_wrap(~country, scales = "free_y") +  # Split by country
  labs(
    title = "Percentage of Non-Citizens Over Time",
    x = "Year",
    y = "Percentage of People",
    color = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"  # Move the legend below the graph
  ) +
  scale_color_manual(
    values = c(
      "Percentage of Non-Citizens" = "red" 
    )
  )
```

### *Average Share of Non-Citizens Over Time*
```{r, echo=FALSE}
merged_data <- merged_data %>%
  mutate(percent_non_citizens = (population_non_citizens / population_total) * 100)

average_non_citizens <- merged_data %>%
  group_by(year) %>%
  summarize(avg_percent_non_citizens = mean(percent_non_citizens, na.rm = TRUE))

ggplot(average_non_citizens, aes(x = year, y = avg_percent_non_citizens)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 2) +
  labs(
    title = "Average Percentage of Non-Citizens Across All Countries",
    x = "Year",
    y = "Average Percentage of Non-Citizens"
  ) +
  theme_minimal()
```

```{r, include=FALSE}
## THIS IS JUST TO FIND THAT THE AVERAGE GROWTH RATE PER YEAR OF THE SHARE OF NON-CITIZENS IS 4.47% PER YEAR OVER ALL COUNTRIES ##
merged_data <- merged_data %>%
  mutate(share_non_citizens = population_non_citizens / population_total)

merged_data <- merged_data %>%
  group_by(country) %>%
  arrange(year) %>%  # Ensure data is sorted by year
  mutate(growth_rate = (share_non_citizens - lag(share_non_citizens)) / lag(share_non_citizens)) %>%
  ungroup()  # Remove grouping by country

overall_average_growth_rate <- merged_data %>%
  summarize(avg_growth_rate = mean(growth_rate, na.rm = TRUE))


overall_average_growth_rate
```

To gain a deeper understanding of our main explanatory variable of interest, we analyzed the percentage of non-citizens in selected countries—Austria, Bulgaria, France, and Germany—alongside the average percentage of non-citizens across the entire sample. For the selected countries, a clear upward trend in the share of non-citizens is evident between 2008 and 2022, although the levels vary significantly between countries. Austria, for instance, stands out with a remarkably high share of non-citizens, reaching approximately 17% in 2022. In contrast, Bulgaria deviates from this general trend, experiencing a decline in the percentage of non-citizens from 1.6% to 1% between 2021 and 2022. 

When we look at the entire sample, the average growth rate of the percentage of non-citizens per year is an impressive 4.47%, highlighting a consistent and significant upward trend. While these findings are descriptive and do not establish causality, they are crucial for contextualizing the broader analysis. They reveal that, on average, the number of non-citizens has grown disproportionately compared to the number of citizens across the countries in our dataset. This disproportionate increase suggests that non-citizens are likely composed of a demographic profile that differs fundamentally from that of citizens. Specifically, the influx of new immigrants tends to be disproportionately male and younger, traits associated with a demographic that is statistically more likely to be suspected or convicted of crime.

This observation is particularly important for understanding the broader context of our study. It offers a plausible explanation for the higher percentage of non-citizens suspected or convicted of crime compared to citizens. By acknowledging the demographic dynamics driving these differences, we can better interpret the patterns observed in the data. Before proceeding with any causal analysis, it is essential to highlight these trends, as they provide a foundational understanding of the structural and demographic factors at play. These findings underscore the importance of considering how the characteristics of non-citizen populations may influence the overall patterns of suspicion and conviction rates observed in the data.

## *Decomposing the Effect of Immigration on Crime*
### *The effect of the share of non-citizens on crime*

### *The effect of immigration on crime while accounting for ideology* (maybe include lagged ideology? the ideology yesterday affects crime today?)

```{r}
## lm model, include right_ideology to the regression
```


### *The effect of immigration on crime while accounting for political polarization*
```{r}
## lm model, add dalton instead of right ideology, then both dalton and right ideology, etc....
## 889f80a398071f900808d11f7180c7ef29d09f9e
```

### *The effect of immigration on crime while accounting for political rhetoric*

```{r}
## include per101 to the regression, etc..
```

SOME YAPYAP

## *Robustness checks*
```{r}
## Chi sqr test, etc....
```

MORE YAP YAP

# Conclusion and discussion
YAPYAPYAP